### 2016-06-14 draft1.tex
now with actual material

              
\documentclass[fleqn,10pt]{wlscirep}
\title{Scientific Reports Title to see here (20 words or less)}

\author[1,*]{Alice Author}
\author[2]{Bob Author}
\author[1,2,+]{Christine Author}
\author[2,+]{Derek Author}
\affil[1]{Affiliation, department, city, postcode, country}
\affil[2]{Affiliation, department, city, postcode, country}

\affil[*]{corresponding.author@email.example}

\affil[+]{these authors contributed equally to this work}

%\keywords{Keyword1, Keyword2, Keyword3}

\begin{abstract}
Example Abstract. Abstract must be under 200 words and not include subheadings or citations. Example Abstract. Abstract must be under 200 words and not include subheadings or citations. Example Abstract. Abstract must be under 200 words and not include subheadings or citations. Example Abstract. Abstract must be under 200 words and not include subheadings or citations. Example Abstract. Abstract must be under 200 words and not include subheadings or citations. Example Abstract. Abstract must be under 200 words and not include subheadings or citations. Example Abstract. Abstract must be under 200 words and not include subheadings or citations. Example Abstract. Abstract must be under 200 words and not include subheadings or citations. 
\end{abstract}
\begin{document}

\flushbottom
\maketitle
% * <john.hammersley@gmail.com> 2015-02-09T12:07:31.197Z:
%
%  Click the title above to edit the author information and abstract
%
\thispagestyle{empty}

\noindent Please note: Abbreviations should be introduced at the first mention in the main text – no abbreviations lists. Suggested structure of main text (not enforced) is provided below.

\section*{Introduction}

The rate at which a heart beats is determined, in part, by the sympathetic and parasympathetic portions of the nervous system. When the sympathetic nervous system is stimulated, it releases epinephrine and norepinephrine  which increase heart rate. The parasympathetic system releases acetylcholine which decreases heart rate.  These two systems, often described as an accelerator and a brake, are not totally independent on each other, rather, they interact through second messengers cAMP and cGMP. \cite{olshansky2008parasympathetic}
Heart rate is also controlled by the baroreflex system. The baroreflex system consists of baroreceptors, tension sensitive nerve endings found in the circulatory system. \cite{ottesen1997modelling} When they sense a change in pressure, they cause a change in the frequency of nerve activity. When pressure (and stretch) rapidly increase, so does the baroreceptor firing rate. \cite{negative1999reflexes} This effects of this signal are not instantaneous, rather, there is a time delay on the order of seconds before the sympathetic and parasympathetic nervous systems respond. \cite{ottesen1997modelling}
Olufsen and Ottesen have developed models of heart rate based on blood pressure measurements.\cite{olufsen2013practical} These models use the baroreflex system and the concentrations of acetylcholine and to predict heart rate.

We used the MIMIC II Waveform Database as the source of the pressure and heart rate data used to test this model.\cite{saeed2011multiparameter} MIMIC contains de-identified data from patients who visited the Beth Israel Deaconess Medical Center ICU. This database contains a wealth of information: diagnoses, demographics, clinical data, such as heart rate and intravenous medications as well as laboratory results.  
A subset of the MIMIC II Waveform Database has been linked to patients found in the MIMIC II Database, allowing us to combine the waveforms with demographic information. From this subset, we selected patients that had numerics records with data recorded once per second with more than five data points available in the first 10 minutes of recorded data, giving us 273 patients meeting this criteria.

\section*{Results}

\subsection*{One Dimensional Optimization of Parameters}
We mimized averaged the mean squared error per patient using the Nelder Mead algorithm as implimented in Julia package Optim. We forced all of the parameters to be non-negative as to be biologically correct.

\subsection*{Clustering}
We clustered the patients based on their age, average heart rate, and SAPS (Simplified Acute Physiology) score, a measure that estimates a patient's risk of death within an intensive care unit.\cite{le1993new} We used the patient's average SAPS since some patients had mutliple admissions resulting in more than one blood pressure-heart rate track.

Through the use of the Clustering Julia package, we created up to 26 clusters based on these variables, and scored each cluster with the sum of its silhouettes, where a higher score means that each member of the cluster is more similar to the other members of the clusters.\cite{rousseeuw1987silhouettes} We found that grouping the patients into two clusters gave the highest score and therefore the best clustering. The sum of silouettes as a function of number of clusters is shown in \ref{fig:sumofsils}, and the patients by cluster in \ref{fig:clusteredpatients}. Cluster 1 patients (n = 165), shown in white, tend to have a lower average heart rate (73 BPM compared to 88 BPM), and be older than patients in cluster 2 (n = 108) (69 years vs 59 years, on average).

\subsection*{Multidimensional Optimization}
We used the Julia language POETs package, which combines simulated anealing with Pareto optimality to generate families of best parameters.\cite{bassen2016jupoets} Using the two clusters formed by k-means, we minimized the averaged mean squared error to create these parameter families. We utilized a monotonically decreasing cooling schedule with five iterations at each temperature. The trade off curve generated using $\alpha$ = .9 is shown in \ref{fig:tradeoffcurve}, and with $\alpha$ =.5  in \ref{fig:tradeoffcurveslower}. To decrease the time neccessary to perform the simulated annealing, we utilized Julia's transparent parallization capabilities. With the @parallel (+) operator, we were able to calculate patient's mean squarred errors in parallel. The speed up from parallization is shown in \ref{fig:executiontime}.

We then selected the ten best sets of parameters from each cluster to examine the performance of the model with the new parameters. A sample patient from cluster 1 is shown in \ref{fig:cluster1patientmulti}, and from cluster 2 in \ref{fig:cluster2patientmulti}.

\subsection*{Sensitivity Using Finite Differences}

The derivitivies of all parameters were estimated using central differences. 
\begin{equation}
\frac{dh}{d\theta_j} = \frac{h(\theta_0+\frac{e_j}{2})-h(\theta_0-\frac{e_j}{2})}{\delta}
\end{equation}
where $\delta=10^{-8}*\theta_j$ and $e_j$ is a vector of length $\delta$ in the $j^{th}$ direction.
To collapse the time dimension, we calculated overall state sensivity coefficents. \cite{stelling2004robustness}
\begin{equation}
S_{0j}(t) = \frac{1}{n_s}p_j \left(\sum_{k=1}^{n_t} \sum_{i=1}^{n_s} \left[ \frac{1}{x_i} \frac{dx_i(t_k)}{dp_j} \right]^2 \right)^{1/2}
\end{equation}
where $n_s$ = 1, as $h$ is the only state varaiable and $n_t$ is the number of time points available for that patient.
From the original values provided by Olufsen and Ottesen, we found that $h_0$ was the most sensitive parameter, followed by $N$, $m_{ach}$, and $m_{nor}$ in that order. Of the four most sensitive equations, three of them ($h_0$, $m_{ach}$ and $m_{nor}$) appear directly in the equation for heart rate, and $N$ indirectly appears, as both $c_{ach}$ and $c_{nor}$ are functions of $N$.
The four parameters that were the most sensitive in our study were among the five most sensitive parameters found by Olufsen and Ottesen, as seen in \ref{tab:OSSCsoriginalParams}.

With the new best parameters, the same four parameters are the most sensitive, as shown in \ref{tab:clusterSensitivity}.

For the multiobjective case, we used the same finite differences, but averaged them not only over the patients, but over the families of parameters-the ten best for each cluster. As with the single objective case, $h_0$ is the most sensitive parameter, followed by $N$, as seen in \ref{fig:MultiobjectiveSensitivityGraph}.

\section*{Discussion}

The Discussion should be succinct and must not contain subheadings.

\section*{Methods}

All calculations were carried out Ubunutu 16.04 using Julia version .0.4.5, with 7.7 GB of available RAM on a Intel Core i7-6700 CPU @ 3.40GHz. Differential equations were solved using the ODE package, with solvers ode23 and ode78, with an absolute and relative tolerance of $10^{-8}$.

\bibliography{draft1}

\section*{Acknowledgements (not compulsory)}

Acknowledgements should be brief, and should not include thanks to anonymous referees and editors, or effusive comments. Grant or contribution numbers may be acknowledged.

\section*{Author contributions statement}

Must include all authors, identified by initials, for example:
A.A. conceived the experiment(s),  A.A. and B.A. conducted the experiment(s), C.A. and D.A. analysed the results.  All authors reviewed the manuscript.

\section*{Additional information}

To include, in this order: \textbf{Accession codes} (where applicable); \textbf{Competing financial interests} (mandatory statement).

The corresponding author is responsible for submitting a \href{http://www.nature.com/srep/policies/index.html#competing}{competing financial interests statement} on behalf of all authors of the paper. This statement must be included in the submitted article file.
\begin{figure}[ht]
		\centering
	\includegraphics[width=\linewidth]{figures/2clusters}
       \caption{The patients, clustered into two groups.}
       \label{fig:clusteredpatients}
\end{figure}

\begin{figure}[ht]
		\centering
	\includegraphics[width=\linewidth]{figures/timedatausingdifferentparamsetsforpatients28611-3018-03-06-18-24n}
       \caption{Performance of the model on a patient in cluster 1. The x represent the true heart rate, the black dots are the mean model prediction from the family of best parameter sets, and the grey gives a 95\% confidence interval.}
       \label{fig:cluster1patientmulti}
\end{figure}

\begin{figure}[ht]
		\centering
	\includegraphics[width=\linewidth]{figures/timedatausingdifferentparamsetsforpatients32582-3106-02-22-22-22n}
       \caption{Performance of the model on a patient in cluster 2. The x represent the true heart rate, the black dots are the mean model prediction from the family of best parameter sets, and the grey gives a 95\% confidence interval.}
       \label{fig:cluster2patientmulti}
\end{figure}

\begin{figure}[ht]
		\centering
       \includegraphics[width=\linewidth]{figures/sumofsilhouettesasafunctionofclustersjustAgeSAPSavgHR}
       \caption{Sum of silhouettes as function of number of clusters}
       \label{fig:sumofsils}
\end{figure}

\begin{figure}[ht]
       \centering
       \includegraphics[width=\linewidth]{figures/runtimecomparison}
       \caption{Switching from serial to parallel computation resulted in a significant speed up. In parallel operation, six cores were used.}
       	\label{fig:MultiobjectiveSensitivityGraph}
\end{figure}

\begin{figure}[ht]
       \centering
       \includegraphics[width=\linewidth]{figures/MultiobjectiveSenstivityGraph}
       \caption{The black circles are from cluster 1, the grey x's represent cluster 2.}
       	\label{fig:MultiobjectiveSensitivityGraph}
\end{figure}

\begin{figure}[ht]
       \centering
       \includegraphics[width=\linewidth]{figures/2runsmultiobjective}
       \caption{The black dots represent the rank 0, or best parameter sets, and the grey dots represent rank 1-4 parameter sets.}
       \label{fig:tradeoffcurve}
\end{figure}

\begin{figure}[ht]
       \centering
       \includegraphics[width=\linewidth]{figures/fastercoolingtradeoff}
       \caption{The black dots represent the rank 0, or best parameter sets, and the grey dots represent rank 1-4 parameter sets.}
       \label{fig:tradeoffcurveslower}
\end{figure}

\begin{table}[h]
\centering
\caption{Average Overall State Sensitivity Coefficents Using Original Parameters}
\label{tab:OSSCsoriginalParams}
\begin{tabular}{lllllllll}
$\frac{dh}{dN}$    & $\frac{dh}{dk_1}$  & $\frac{dh}{d\tau_1}$ & $\frac{dh}{d\tau_2}$ &$\frac{dh}{d\tau_{ach}}$ & $\frac{dh}{d\tau_{nor}}$ & $\frac{dh}{dh_0}$   & $\frac{dh}{dm_{nor}}$ & $\frac{dh}{dm_{ach}}$ \\
15.3620 & 0.0792 & 0.0740  & 0.0156  & 0.0540   & 0.0006    & 19.9294 & 0.5740  & 14.5868
\end{tabular}
\end{table}

\begin{table}[]
\centering
\caption{Sensitivity by clusters}
\label{tab:clusterSensitivity}
\begin{tabular}{|l|l|l|l|l|l|l|l|l|l|}
\hline
Cluster   & $\frac{dh}{dN}$ & $\frac{dh}{k_1}$ & $\frac{dh}{\tau_1}$ & $\frac{dh}{tau_2}$ & $\frac{dh}{\tau_{ach}}$ & $\frac{dh}{\tau_{nor}}$ & $\frac{dh}{h0}$ & $\frac{dh}{m_{nor}}$ & $\frac{dh}{m_{ach}}$ \\ \hline
Overall   & 4.8352          & 0.2500           & 0.5088              & 0.0033             & 0.0204                  & 0.5720                  & 19.9294         & 1.1339               & 4.4651               \\ \hline
cluster 1 & 4.7152          & 0.2456           & 0.5027              & 0.0033             & 0.0201                  & 0.5755                  & 19.4780         & 1.1438               & 4.3593               \\ \hline
cluster 2 & 4.9137          & 0.2528           & 0.5128              & 0.0033             & 0.0206                  & 0.5697                  & 20.2248         & 1.1275               & 4.5344               \\ \hline
\end{tabular}
\end{table}

\end{document}

### 2016-06-14 formattemplate.tex
now with actual material

--- 2016-06-14 draft1.tex
@@
 
-The rate at which a heart beats is determined, in part, by the sympathetic and parasympathetic portions of the nervous system.
-When the sympathetic nervous system is stimulated, it releases epinephrine and norepinephrine  which increase heart rate.
-The parasympathetic system releases acetylcholine which decreases heart rate.
-These two systems, often described as an accelerator and a brake, are not totally independent on each other, rather, they interact through second messengers cAMP and cGMP.
-\cite{olshansky2008parasympathetic}
-Heart rate is also controlled by the baroreflex system.
-The baroreflex system consists of baroreceptors, tension sensitive nerve endings found in the circulatory system.
-\cite{ottesen1997modelling} When they sense a change in pressure, they cause a change in the frequency of nerve activity.
-When pressure (and stretch) rapidly increase, so does the baroreceptor firing rate.
-\cite{negative1999reflexes} This effects of this signal are not instantaneous, rather, there is a time delay on the order of seconds before the sympathetic and parasympathetic nervous systems respond.
-\cite{ottesen1997modelling}
-Olufsen and Ottesen have developed models of heart rate based on blood pressure measurements.\cite{olufsen2013practical} These models use the baroreflex system and the concentrations of acetylcholine and to predict heart rate.
-
-We used the MIMIC II Waveform Database as the source of the pressure and heart rate data used to test this model.\cite{saeed2011multiparameter} MIMIC contains de-identified data from patients who visited the Beth Israel Deaconess Medical Center ICU.
-This database contains a wealth of information: diagnoses, demographics, clinical data, such as heart rate and intravenous medications as well as laboratory results.
-A subset of the MIMIC II Waveform Database has been linked to patients found in the MIMIC II Database, allowing us to combine the waveforms with demographic information.
-From this subset, we selected patients that had numerics records with data recorded once per second with more than five data points available in the first 10 minutes of recorded data, giving us 273 patients meeting this criteria.
+The Introduction section, of referenced text\cite{Figueredo:2009dg} expands on the background of the work (some overlap with the Abstract is acceptable).
+The introduction should not include subheadings.
 
@@
 
-\subsection*{One Dimensional Optimization of Parameters}
-We mimized averaged the mean squared error per patient using the Nelder Mead algorithm as implimented in Julia package Optim.
-We forced all of the parameters to be non-negative as to be biologically correct.
+Up to three levels of \textbf{subheading} are permitted. Subheadings should not be numbered.
 
-\subsection*{Clustering}
-We clustered the patients based on their age, average heart rate, and SAPS (Simplified Acute Physiology) score, a measure that estimates a patient's risk of death within an intensive care unit.\cite{le1993new} We used the patient's average SAPS since some patients had mutliple admissions resulting in more than one blood pressure-heart rate track.
+\subsection*{Subsection}
 
-Through the use of the Clustering Julia package, we created up to 26 clusters based on these variables, and scored each cluster with the sum of its silhouettes, where a higher score means that each member of the cluster is more similar to the other members of the clusters.\cite{rousseeuw1987silhouettes} We found that grouping the patients into two clusters gave the highest score and therefore the best clustering.
-The sum of silouettes as a function of number of clusters is shown in \ref{fig:sumofsils}, and the patients by cluster in \ref{fig:clusteredpatients}.
-Cluster 1 patients (n = 165), shown in white, tend to have a lower average heart rate (73 BPM compared to 88 BPM), and be older than patients in cluster 2 (n = 108) (69 years vs 59 years, on average).
+Example text under a subsection. Bulleted lists may be used where appropriate, e.g.
 
-\subsection*{Multidimensional Optimization}
-We used the Julia language POETs package, which combines simulated anealing with Pareto optimality to generate families of best parameters.\cite{bassen2016jupoets} Using the two clusters formed by k-means, we minimized the averaged mean squared error to create these parameter families.
-We utilized a monotonically decreasing cooling schedule with five iterations at each temperature.
-The trade off curve generated using $\alpha$ = .9 is shown in \ref{fig:tradeoffcurve}, and with $\alpha$ =.5  in \ref{fig:tradeoffcurveslower}.
-To decrease the time neccessary to perform the simulated annealing, we utilized Julia's transparent parallization capabilities.
-With the @parallel (+) operator, we were able to calculate patient's mean squarred errors in parallel.
-The speed up from parallization is shown in \ref{fig:executiontime}.
+\begin{itemize}
+\item First item
+\item Second item
+\end{itemize}
 
-We then selected the ten best sets of parameters from each cluster to examine the performance of the model with the new parameters.
-A sample patient from cluster 1 is shown in \ref{fig:cluster1patientmulti}, and from cluster 2 in \ref{fig:cluster2patientmulti}.
+\subsubsection*{Third-level section}
 
-\subsection*{Sensitivity Using Finite Differences}
-
-The derivitivies of all parameters were estimated using central differences. 
-\begin{equation}
-\frac{dh}{d\theta_j} = \frac{h(\theta_0+\frac{e_j}{2})-h(\theta_0-\frac{e_j}{2})}{\delta}
-\end{equation}
-where $\delta=10^{-8}*\theta_j$ and $e_j$ is a vector of length $\delta$ in the $j^{th}$ direction.
-To collapse the time dimension, we calculated overall state sensivity coefficents. \cite{stelling2004robustness}
-\begin{equation}
-S_{0j}(t) = \frac{1}{n_s}p_j \left(\sum_{k=1}^{n_t} \sum_{i=1}^{n_s} \left[ \frac{1}{x_i} \frac{dx_i(t_k)}{dp_j} \right]^2 \right)^{1/2}
-\end{equation}
-where $n_s$ = 1, as $h$ is the only state varaiable and $n_t$ is the number of time points available for that patient.
-From the original values provided by Olufsen and Ottesen, we found that $h_0$ was the most sensitive parameter, followed by $N$, $m_{ach}$, and $m_{nor}$ in that order.
-Of the four most sensitive equations, three of them ($h_0$, $m_{ach}$ and $m_{nor}$) appear directly in the equation for heart rate, and $N$ indirectly appears, as both $c_{ach}$ and $c_{nor}$ are functions of $N$.
-The four parameters that were the most sensitive in our study were among the five most sensitive parameters found by Olufsen and Ottesen, as seen in \ref{tab:OSSCsoriginalParams}.
-
-With the new best parameters, the same four parameters are the most sensitive, as shown in \ref{tab:clusterSensitivity}.
-
-For the multiobjective case, we used the same finite differences, but averaged them not only over the patients, but over the families of parameters-the ten best for each cluster.
-As with the single objective case, $h_0$ is the most sensitive parameter, followed by $N$, as seen in \ref{fig:MultiobjectiveSensitivityGraph}.
+Topical subheadings are allowed.
 
@@
 
-All calculations were carried out Ubunutu 16.04 using Julia version .0.4.5, with 7.7 GB of available RAM on a Intel Core i7-6700 CPU @ 3.40GHz.
-Differential equations were solved using the ODE package, with solvers ode23 and ode78, with an absolute and relative tolerance of $10^{-8}$.
+Topical subheadings are allowed. Authors must ensure that their Methods section includes adequate experimental and characterization data necessary for others in the field to reproduce their work.
 
-\bibliography{draft1}
+\bibliography{sample}
+
+\noindent LaTeX formats citations and references automatically using the bibliography records in your .bib file, which you can edit via the project menu.
+Use the cite command for an inline citation, e.g.
+\cite{Figueredo:2009dg}.
 
@@
 This statement must be included in the submitted article file.
-\begin{figure}[ht]
-		\centering
-	\includegraphics[width=\linewidth]{figures/2clusters}
-       \caption{The patients, clustered into two groups.}
-       \label{fig:clusteredpatients}
-\end{figure}
 
@@
 \centering
-	\includegraphics[width=\linewidth]{figures/timedatausingdifferentparamsetsforpatients28611-3018-03-06-18-24n}
-       \caption{Performance of the model on a patient in cluster 1.
-The x represent the true heart rate, the black dots are the mean model prediction from the family of best parameter sets, and the grey gives a 95\% confidence interval.}
-       \label{fig:cluster1patientmulti}
+\includegraphics[width=\linewidth]{stream}
+\caption{Legend (350 words max). Example legend text.}
+\label{fig:stream}
 \end{figure}
 
-\begin{figure}[ht]
+\begin{table}[ht]
 \centering
-	\includegraphics[width=\linewidth]{figures/timedatausingdifferentparamsetsforpatients32582-3106-02-22-22-22n}
-       \caption{Performance of the model on a patient in cluster 2.
-The x represent the true heart rate, the black dots are the mean model prediction from the family of best parameter sets, and the grey gives a 95\% confidence interval.}
-       \label{fig:cluster2patientmulti}
-\end{figure}
-
-\begin{figure}[ht]
-		\centering
-       \includegraphics[width=\linewidth]{figures/sumofsilhouettesasafunctionofclustersjustAgeSAPSavgHR}
-       \caption{Sum of silhouettes as function of number of clusters}
-       \label{fig:sumofsils}
-\end{figure}
-
-\begin{figure}[ht]
-       \centering
-       \includegraphics[width=\linewidth]{figures/runtimecomparison}
-       \caption{Switching from serial to parallel computation resulted in a significant speed up. In parallel operation, six cores were used.}
-       	\label{fig:MultiobjectiveSensitivityGraph}
-\end{figure}
-
-\begin{figure}[ht]
-       \centering
-       \includegraphics[width=\linewidth]{figures/MultiobjectiveSenstivityGraph}
-       \caption{The black circles are from cluster 1, the grey x's represent cluster 2.}
-       	\label{fig:MultiobjectiveSensitivityGraph}
-\end{figure}
-
-\begin{figure}[ht]
-       \centering
-       \includegraphics[width=\linewidth]{figures/2runsmultiobjective}
-       \caption{The black dots represent the rank 0, or best parameter sets, and the grey dots represent rank 1-4 parameter sets.}
-       \label{fig:tradeoffcurve}
-\end{figure}
-
-\begin{figure}[ht]
-       \centering
-       \includegraphics[width=\linewidth]{figures/fastercoolingtradeoff}
-       \caption{The black dots represent the rank 0, or best parameter sets, and the grey dots represent rank 1-4 parameter sets.}
-       \label{fig:tradeoffcurveslower}
-\end{figure}
-
-\begin{table}[h]
-\centering
-\caption{Average Overall State Sensitivity Coefficents Using Original Parameters}
-\label{tab:OSSCsoriginalParams}
-\begin{tabular}{lllllllll}
-$\frac{dh}{dN}$    & $\frac{dh}{dk_1}$  & $\frac{dh}{d\tau_1}$ & $\frac{dh}{d\tau_2}$ &$\frac{dh}{d\tau_{ach}}$ & $\frac{dh}{d\tau_{nor}}$ & $\frac{dh}{dh_0}$   & $\frac{dh}{dm_{nor}}$ & $\frac{dh}{dm_{ach}}$ \\
-15.3620 & 0.0792 & 0.0740  & 0.0156  & 0.0540   & 0.0006    & 19.9294 & 0.5740  & 14.5868
+\begin{tabular}{|l|l|l|}
+\hline
+Condition & n & p \\
+\hline
+A & 5 & 0.1 \\
+\hline
+B & 10 & 0.01 \\
+\hline
 \end{tabular}
+\caption{\label{tab:example}Legend (350 words max). Example legend text.}
 \end{table}
 
-\begin{table}[]
-\centering
-\caption{Sensitivity by clusters}
-\label{tab:clusterSensitivity}
-\begin{tabular}{|l|l|l|l|l|l|l|l|l|l|}
-\hline
-Cluster   & $\frac{dh}{dN}$ & $\frac{dh}{k_1}$ & $\frac{dh}{\tau_1}$ & $\frac{dh}{tau_2}$ & $\frac{dh}{\tau_{ach}}$ & $\frac{dh}{\tau_{nor}}$ & $\frac{dh}{h0}$ & $\frac{dh}{m_{nor}}$ & $\frac{dh}{m_{ach}}$ \\ \hline
-Overall   & 4.8352          & 0.2500           & 0.5088              & 0.0033             & 0.0204                  & 0.5720                  & 19.9294         & 1.1339               & 4.4651               \\ \hline
-cluster 1 & 4.7152          & 0.2456           & 0.5027              & 0.0033             & 0.0201                  & 0.5755                  & 19.4780         & 1.1438               & 4.3593               \\ \hline
-cluster 2 & 4.9137          & 0.2528           & 0.5128              & 0.0033             & 0.0206                  & 0.5697                  & 20.2248         & 1.1275               & 4.5344               \\ \hline
-\end{tabular}
-\end{table}
+Figures and tables can be referenced in LaTeX using the ref command, e.g. Figure \ref{fig:stream} and Table \ref{tab:example}.
 

### 2016-06-14 draft1.tex
small changes

@@
 
-\author[1,*]{Alice Author}
-\author[2]{Bob Author}
-\author[1,2,+]{Christine Author}
-\author[2,+]{Derek Author}
-\affil[1]{Affiliation, department, city, postcode, country}
-\affil[2]{Affiliation, department, city, postcode, country}
+\author[1]{Rachel LeCover}
+\author[1,*]{Jeffrey D. Varner}
+\affil[1]{Affiliation, School of Chemical and Biomolecular Engineering, Ithaca NY, 14852, USA}
 
-\affil[*]{corresponding.author@email.example}
-
-\affil[+]{these authors contributed equally to this work}
+\affil[*]{jdv27@cornell.edu}
 
@@
 
+To our knowledge, this is the first study to attempt to model heart rate as a function of blood pressure for such a large number of patients.
+Additionally, many previous studies were performed with healthy patients, in contrast, this study focuses on intensive care unit patients.
+
 \section*{Results}
@@
 The trade off curve generated using $\alpha$ = .9 is shown in \ref{fig:tradeoffcurve}, and with $\alpha$ =.5  in \ref{fig:tradeoffcurveslower}.
+We used the parameter families from the slower cooling for further analysis.
+
 To decrease the time neccessary to perform the simulated annealing, we utilized Julia's transparent parallization capabilities.

### 2016-06-15 draft1.tex
added to 1d opt

@@
-              
 \documentclass[fleqn,10pt]{wlscirep}
 \title{Scientific Reports Title to see here (20 words or less)}
-
 \author[1]{Rachel LeCover}
@@
 \affil[1]{Affiliation, School of Chemical and Biomolecular Engineering, Ithaca NY, 14852, USA}
-
 \affil[*]{jdv27@cornell.edu}
-
 %\keywords{Keyword1, Keyword2, Keyword3}
-
 \begin{abstract}
@@
 \begin{document}
-
 \flushbottom
@@
 \thispagestyle{empty}
-
 \noindent Please note: Abbreviations should be introduced at the first mention in the main text – no abbreviations lists. Suggested structure of main text (not enforced) is provided below.
-
 \section*{Introduction}
-
 The rate at which a heart beats is determined, in part, by the sympathetic and parasympathetic portions of the nervous system.
@@
 Olufsen and Ottesen have developed models of heart rate based on blood pressure measurements.\cite{olufsen2013practical} These models use the baroreflex system and the concentrations of acetylcholine and to predict heart rate.
-
 We used the MIMIC II Waveform Database as the source of the pressure and heart rate data used to test this model.\cite{saeed2011multiparameter} MIMIC contains de-identified data from patients who visited the Beth Israel Deaconess Medical Center ICU.
@@
 From this subset, we selected patients that had numerics records with data recorded once per second with more than five data points available in the first 10 minutes of recorded data, giving us 273 patients meeting this criteria.
-
 To our knowledge, this is the first study to attempt to model heart rate as a function of blood pressure for such a large number of patients.
 Additionally, many previous studies were performed with healthy patients, in contrast, this study focuses on intensive care unit patients.
-
 \section*{Results}
-
 \subsection*{One Dimensional Optimization of Parameters}
-We mimized averaged the mean squared error per patient using the Nelder Mead algorithm as implimented in Julia package Optim.
-We forced all of the parameters to be non-negative as to be biologically correct.
-
+We minimized averaged the mean squared error per patient using the Nelder Mead algorithm as implemented in Julia package Optim.
+We forced all of the parameters to be non-negative as to be biologically correct.The original and estimated values of the parameters are shown in \ref{tab:SingleParams}.
+The value of $\tau_2$ decreased dramatically in the estimated parameter set, increasing the effect of $n_1$ on the firing rate.
+Additionally, $\tau_{ach}$ increased significantly, reducing the changes in $c_{ach}$.
+$k_1$ and $\beta$ were held constant as to reduce the search space.
+$\alpha$, the parameter used to smooth the pressure data was held constant at 1.5 to smooth the data with a minimal phase shift.
+The maximum firing rate, $M$, was held constant, as it only appears in this system of equations by normalizing $n$.
+The sympathetic nervous system delay, $\tau_d$, was held constant, owing to the propagation of discontinuities.
+\cite{baker1997pitfalls}
 \subsection*{Clustering}
-We clustered the patients based on their age, average heart rate, and SAPS (Simplified Acute Physiology) score, a measure that estimates a patient's risk of death within an intensive care unit.\cite{le1993new} We used the patient's average SAPS since some patients had mutliple admissions resulting in more than one blood pressure-heart rate track.
-
+We clustered the patients based on their age, average heart rate, and SAPS (Simplified Acute Physiology) score, a measure that estimates a patient's risk of death within an intensive care unit.\cite{le1993new} We used the patient's average SAPS since some patients had multiple admissions resulting in more than one blood pressure-heart rate track.
 Through the use of the Clustering Julia package, we created up to 26 clusters based on these variables, and scored each cluster with the sum of its silhouettes, where a higher score means that each member of the cluster is more similar to the other members of the clusters.\cite{rousseeuw1987silhouettes} We found that grouping the patients into two clusters gave the highest score and therefore the best clustering.
-The sum of silouettes as a function of number of clusters is shown in \ref{fig:sumofsils}, and the patients by cluster in \ref{fig:clusteredpatients}.
+The sum of silhouettes as a function of number of clusters is shown in \ref{fig:sumofsils}, and the patients by cluster in \ref{fig:clusteredpatients}.
 Cluster 1 patients (n = 165), shown in white, tend to have a lower average heart rate (73 BPM compared to 88 BPM), and be older than patients in cluster 2 (n = 108) (69 years vs 59 years, on average).
-
 \subsection*{Multidimensional Optimization}
@@
 We used the parameter families from the slower cooling for further analysis.
-
-To decrease the time neccessary to perform the simulated annealing, we utilized Julia's transparent parallization capabilities.
+To decrease the time necessary to perform the simulated annealing, we utilized Julia's transparent parallelization capabilities.
 With the @parallel (+) operator, we were able to calculate patient's mean squarred errors in parallel.
-The speed up from parallization is shown in \ref{fig:executiontime}.
-
+The speed up from parallelization is shown in \ref{fig:executiontime}.
 We then selected the ten best sets of parameters from each cluster to examine the performance of the model with the new parameters.
@@
 \subsection*{Sensitivity Using Finite Differences}
-
-The derivitivies of all parameters were estimated using central differences. 
+The derivatives of all parameters were estimated using central differences. 
 \begin{equation}
@@
 where $\delta=10^{-8}*\theta_j$ and $e_j$ is a vector of length $\delta$ in the $j^{th}$ direction.
-To collapse the time dimension, we calculated overall state sensivity coefficents. \cite{stelling2004robustness}
+To collapse the time dimension, we calculated overall state sensitivity coefficients. \cite{stelling2004robustness}
 \begin{equation}
@@
 \end{equation}
-where $n_s$ = 1, as $h$ is the only state varaiable and $n_t$ is the number of time points available for that patient.
+where $n_s$ = 1, as $h$ is the only state variable and $n_t$ is the number of time points available for that patient.
 From the original values provided by Olufsen and Ottesen, we found that $h_0$ was the most sensitive parameter, followed by $N$, $m_{ach}$, and $m_{nor}$ in that order.
@@
 The four parameters that were the most sensitive in our study were among the five most sensitive parameters found by Olufsen and Ottesen, as seen in \ref{tab:OSSCsoriginalParams}.
-
 With the new best parameters, the same four parameters are the most sensitive, as shown in \ref{tab:clusterSensitivity}. 
-
 For the multiobjective case, we used the same finite differences, but averaged them not only over the patients, but over the families of parameters-the ten best for each cluster.
 As with the single objective case, $h_0$ is the most sensitive parameter, followed by $N$, as seen in \ref{fig:MultiobjectiveSensitivityGraph}.
-
 \section*{Discussion}
-
 The Discussion should be succinct and must not contain subheadings.
-
 \section*{Methods}
-
 All calculations were carried out Ubunutu 16.04 using Julia version .0.4.5, with 7.7 GB of available RAM on a Intel Core i7-6700 CPU @ 3.40GHz.
 Differential equations were solved using the ODE package, with solvers ode23 and ode78, with an absolute and relative tolerance of $10^{-8}$.
-
 \bibliography{draft1}
-
 \section*{Acknowledgements (not compulsory)}
-
 Acknowledgements should be brief, and should not include thanks to anonymous referees and editors, or effusive comments. Grant or contribution numbers may be acknowledged.
-
 \section*{Author contributions statement}
-
 Must include all authors, identified by initials, for example:
 A.A. conceived the experiment(s),  A.A. and B.A. conducted the experiment(s), C.A. and D.A. analysed the results.  All authors reviewed the manuscript. 
-
 \section*{Additional information}
-
 To include, in this order: \textbf{Accession codes} (where applicable); \textbf{Competing financial interests} (mandatory statement). 
-
 The corresponding author is responsible for submitting a \href{http://www.nature.com/srep/policies/index.html#competing}{competing financial interests statement} on behalf of all authors of the paper.
@@
 \end{figure}
-
 \begin{figure}[ht]
@@
 \end{figure} 
-
 \begin{figure}[ht]
@@
 \end{figure}
-
+\begin{table}[ht]
+\centering
+\caption{Original and estimated parameter values. The bolded parameters were held constant.}
+\label{tab:SingleParams}
+\begin{tabular}{|l|l|l|l|l|l|l|l|l|l|l|l|l|l|l|}
+\hline
+Parameter                                 & \textbf{$\alpha$} & $N$    & \textbf{$M$} & $k_1$ & \textbf{$k_2$} & $\tau_1$ & $\tau_2$ & $\tau_{ach}$ & $\tau_{nor}$ & \textbf{$\beta$} & $h_0$ & $m_{nor}$ & $m_{ach}$ & \textbf{$\tau_d$} \\ \hline
+Original Value                            & 1.5               & 75     & 120          & 1.5   & 0.5            & 0.5      & 250      & 0.5          & 0.5          & 6                & 1.67  & 0.96      & 0.7       & 7                 \\ \hline
+Parameter Estimations from MIMIC II  Data & 1.5               & 75.319 & 120          & 2.299 & 0.5            & 1.345    & 0.126    & 250.339      & 3.185        & 6                & 1.412 & 0.160     & 2.35E-06  & 7                 \\ \hline
+\end{tabular}
+\end{table}
 \begin{table}[h]
@@
 \end{table} 
-
 \begin{table}[]
@@
 \end{table}
-
 \end{document}

### 2016-06-16 draft1.tex
changed cooling curves to match

@@
+              
 \documentclass[fleqn,10pt]{wlscirep}
@@
 Olufsen and Ottesen have developed models of heart rate based on blood pressure measurements.\cite{olufsen2013practical} These models use the baroreflex system and the concentrations of acetylcholine and to predict heart rate.
+
 We used the MIMIC II Waveform Database as the source of the pressure and heart rate data used to test this model.\cite{saeed2011multiparameter} MIMIC contains de-identified data from patients who visited the Beth Israel Deaconess Medical Center ICU.
@@
        \caption{Switching from serial to parallel computation resulted in a significant speed up. In parallel operation, six cores were used.}
-        \label{fig:MultiobjectiveSensitivityGraph}
+        \label{fig:executiontime}
 \end{figure}
@@
        \centering
-       \includegraphics[width=\linewidth]{figures/2runsmultiobjective}
+       \includegraphics[width=\linewidth]{figures/TradeOffCurveFasterCooling}
        \caption{The black dots represent the rank 0, or best parameter sets, and the grey dots represent rank 1-4 parameter sets.}
@@
        \centering
-       \includegraphics[width=\linewidth]{figures/fastercoolingtradeoff}
+       \includegraphics[width=\linewidth]{figures/TradeOffCurveSlowerCooling}
        \caption{The black dots represent the rank 0, or best parameter sets, and the grey dots represent rank 1-4 parameter sets.}

### 2016-06-16 draft1.tex
added comparison of original vs estimated params

@@
 Additionally, $\tau_{ach}$ increased significantly, reducing the changes in $c_{ach}$.
+As shown in figure \ref{fig:paramscomparsion}, the estimated parameters tend to smooth out the varaibility in the predicted heart rate compared to the original parameters.
+
 $k_1$ and $\beta$ were held constant as to reduce the search space.
@@
 \cite{baker1997pitfalls}
+
 \subsection*{Clustering}
@@
 This statement must be included in the submitted article file.
+
+\begin{figure}[ht]
+                \centering
+        \includegraphics[width=\linewidth]{figures/paramscomparisons28611-3108-03-06-18-24n}
+       \caption{The black xs are the measured heart rate, the dark gray is the model prediction using the original parameters and the light gray is the model prediction using the original parameters.}
+       \label{fig:paramscomparsion}
+\end{figure}
+
 \begin{figure}[ht]
@@
 \end{figure}
+
 \begin{figure}[ht]

### 2016-06-16 draft1.tex
changed multiobjective senstivity graph

@@
 \cite{ottesen1997modelling}
-Olufsen and Ottesen have developed models of heart rate based on blood pressure measurements.\cite{olufsen2013practical} These models use the baroreflex system and the concentrations of acetylcholine and to predict heart rate.
+Olufsen and Ottesen have developed models of heart rate based on blood pressure measurements.\cite{olufsen2013practical} These models use the baroreflex system and the concentrations of acetylcholine and noradrenaline to predict heart rate.
 
@@
 We minimized averaged the mean squared error per patient using the Nelder Mead algorithm as implemented in Julia package Optim.
-We forced all of the parameters to be non-negative as to be biologically correct.The original and estimated values of the parameters are shown in \ref{tab:SingleParams}.
+We forced all of the parameters to be non-negative as to be biologically correct.The original and estimated values of the parameters are shown in Table \ref{tab:SingleParams}.
 The value of $\tau_2$ decreased dramatically in the estimated parameter set, increasing the effect of $n_1$ on the firing rate.
 Additionally, $\tau_{ach}$ increased significantly, reducing the changes in $c_{ach}$.
-As shown in figure \ref{fig:paramscomparsion}, the estimated parameters tend to smooth out the varaibility in the predicted heart rate compared to the original parameters.
+As shown in Figure \ref{fig:paramscomparsion}, the estimated parameters tend to smooth out the varaibility in the predicted heart rate compared to the original parameters.
+The change in parameters reduced the mean squared error, averaged over all of the selected patients, from 1167 to 349.
 
@@
 A sample patient from cluster 1 is shown in \ref{fig:cluster1patientmulti}, and from cluster 2 in \ref{fig:cluster2patientmulti}.
+The new parameter families reduced the averaged mean squared error even further than the parameters estimated using Nelder Mead-the ten best families for cluster 1 reduced the averaged mean squared error to 135 or less, and for cluster two, the averaged mean squared error was reduced to 178, or less.
 
@@
        \centering
-       \includegraphics[width=\linewidth]{figures/MultiobjectiveSenstivityGraph}
-       \caption{The black circles are from cluster 1, the grey x's represent cluster 2.}
+       \includegraphics[width=\linewidth]{figures/MultiobjectiveSenstivityGraph10familiesofNoErrorBars}
+       \caption{The black circles are from cluster 1, the grey x's represent cluster 2. Error bars are omitted for the sake of clarity.}
         \label{fig:MultiobjectiveSensitivityGraph}

### 2016-06-20 draft1.tex
Changed model description

@@
 \cite{ottesen1997modelling}
-Olufsen and Ottesen have developed models of heart rate based on blood pressure measurements.\cite{olufsen2013practical} These models use the baroreflex system and the concentrations of acetylcholine and noradrenaline to predict heart rate.
+
+Olufsen and Ottesen have developed models of heart rate based on blood pressure measurements.\cite{olufsen2013practical} From the blood pressure measurement, the model predicts a firing rate for different types of receptors, which is then used to predict the response times of the parasympathetic and sympathetic nervous system.\cite{ottesen1996non} The response times are then used to predict dimensionless norepinephrine and acetylcholine concentrations, which are finally used to predict heart rate.
 

### 2016-06-20 draft1.tex
changed 1D optimization to use parameters from NLopt

@@
 \subsection*{One Dimensional Optimization of Parameters}
-We minimized averaged the mean squared error per patient using the Nelder Mead algorithm as implemented in Julia package Optim.
-We forced all of the parameters to be non-negative as to be biologically correct.The original and estimated values of the parameters are shown in Table \ref{tab:SingleParams}.
-The value of $\tau_2$ decreased dramatically in the estimated parameter set, increasing the effect of $n_1$ on the firing rate.
+We minimized averaged the mean squared error per patient using the Nelder Mead algorithm as implemented in Julia package NLopt.
+We forced all of the parameters to be non-negative as to be biologically correct.The original and estimated values of the parameters are shown in Table \ref{tab:SingleParamsNLopt}.
+%The value of $\tau_2$ decreased dramatically in the estimated parameter set, increasing the effect of $n_1$ on the firing rate.
 Additionally, $\tau_{ach}$ increased significantly, reducing the changes in $c_{ach}$.
@@
 The change in parameters reduced the mean squared error, averaged over all of the selected patients, from 1167 to 349.
+
+Within the estimated parameter set, the value of $N$, the baseline firing rate increased, and both $k_1$ and $\tau_1$, representing a change in the response rate of the faster firing baroreceptors in the model.The baseline heart rate ($h_0$) decreased slightly, to 88 bpm, from the initial 100 bpm.
+Both $\tau_{ach}$ and $m_{ach}$, which determine how acetylcholine plays into the model, decreased dramatically in the estimated parameter set, with the reduction in $m_{ach}$ reducing the effects of a change in acetylcholine on heart rate, and the decrease in $\tau_{ach}$ increasing the rate of acetylcholine change.
+The estimated parameter set reduced the averaged mean squared error to 352 from 1167.
+As seen in Figure \ref{fig:paramscomparsionNLopt}, the estimated parameters slightly overestimate heart rate for this patient, but they perform far better than the original parameters, which severely underestimate the heart rate.
 
@@
 
+%\begin{figure}[ht]
+%                \centering
+%        \includegraphics[width=\linewidth]{figures/paramscomparisons28611-3108-03-06-18-24n}
+%       \caption{The black xs are the measured heart rate, the dark gray is the model prediction using the original parameters and the light gray is the model prediction using the original parameters.}
+%       \label{fig:paramscomparsion}
+%\end{figure}
+
 \begin{figure}[ht]
                 \centering
-        \includegraphics[width=\linewidth]{figures/paramscomparisons28611-3108-03-06-18-24n}
+        \includegraphics[width=\linewidth]{figures/s27193-3271-03-15-13-26n}
        \caption{The black xs are the measured heart rate, the dark gray is the model prediction using the original parameters and the light gray is the model prediction using the original parameters.}
-       \label{fig:paramscomparsion}
+       \label{fig:paramscomparsionNLopt}
 \end{figure}
@@
 \end{figure}
+
+%\begin{table}[ht]
+%\centering
+%\caption{Original and estimated parameter values. The bolded parameters were held constant.}
+%\label{tab:SingleParamsOptim}
+%\begin{tabular}{|l|l|l|l|l|l|l|l|l|l|l|l|l|l|l|}
+%\hline
+%Parameter                                 & \textbf{$\alpha$} & $N$    & \textbf{$M$} & $k_1$ & \textbf{$k_2$} & $\tau_1$ & $\tau_2$ & $\tau_{ach}$ & $\tau_{nor}$ & \textbf{$\beta$} & $h_0$ & $m_{nor}$ & $m_{ach}$ & \textbf{$\tau_d$} \\ \hline
+%Original Value                            & 1.5               & 75     & 120          & 1.5   & 0.5            & 0.5      & 250      & 0.5          & 0.5          & 6                & 1.67  & 0.96      & 0.7       & 7                 \\ \hline
+%Parameter Estimations from MIMIC II  Data & 1.5               & 75.319 & 120          & 2.299 & 0.5            & 1.345    & 0.126    & 250.339      & 3.185        & 6                & 1.412 & 0.160     & 2.35E-06  & 7                 \\ \hline
+%\end{tabular}
+%\end{table}
+
 \begin{table}[ht]
@@
 \caption{Original and estimated parameter values. The bolded parameters were held constant.}
-\label{tab:SingleParams}
+\label{tab:SingleParamsNLopt}
 \begin{tabular}{|l|l|l|l|l|l|l|l|l|l|l|l|l|l|l|}
@@
 Parameter                                 & \textbf{$\alpha$} & $N$    & \textbf{$M$} & $k_1$  & \textbf{$k_2$} & $\tau_1$ & $\tau_2$ & $\tau_{ach}$ & $\tau_{nor}$ & \textbf{$\beta$} & $h_0$ & $m_{nor}$ & $m_{ach}$ & \textbf{$\tau_d$} \\ \hline
-Original Value                            & 1.5               & 75     & 120          & 1.5   & 0.5            & 0.5      & 250      & 0.5          & 0.5          & 6                & 1.67  & 0.96      & 0.7       & 7                 \\ \hline
-Parameter Estimations from MIMIC II  Data & 1.5               & 75.319 & 120          & 2.299 & 0.5            & 1.345    & 0.126    & 250.339      & 3.185        & 6                & 1.412 & 0.160     & 2.35E-06  & 7                 \\ \hline
+Original Value                            & \textbf{1.5}      & 75     & \textbf{120} & 1.5    & \textbf{0.5}   & 0.5      & 250      & 0.5          & 0.5          & \textbf{6}       & 1.67  & 0.96      & 0.7       & \textbf{7}        \\ \hline
+Parameter Estimations from MIMIC II  Data & \textbf{1.5}      & 107.33 & \textbf{120} & 41.139 & \textbf{0.5}   & 14.062   & 202.925  & 2.94E-05     & 8.156        & \textbf{6}       & 1.474 & 0.165     & 5.20E-02  & \textbf{7}        \\ \hline
 \end{tabular}
 \end{table}
+
 \begin{table}[h]

### 2016-06-21 draft1.tex
Added information about barroreceptors

@@
 
-Olufsen and Ottesen have developed models of heart rate based on blood pressure measurements.\cite{olufsen2013practical} From the blood pressure measurement, the model predicts a firing rate for different types of receptors, which is then used to predict the response times of the parasympathetic and sympathetic nervous system.\cite{ottesen1996non} The response times are then used to predict dimensionless norepinephrine and acetylcholine concentrations, which are finally used to predict heart rate.
+Olufsen and Ottesen have developed models of heart rate based on blood pressure measurements.\cite{olufsen2013practical} From the blood pressure measurement, the model predicts a firing rate for different types of receptors, which is then used to predict the response times of the parasympathetic and sympathetic nervous system.
+The response times are then used to predict dimensionless norepinephrine and acetylcholine concentrations, which are finally used to predict heart rate.
 
@@
 
-Within the estimated parameter set, the value of $N$, the baseline firing rate increased, and both $k_1$ and $\tau_1$, representing a change in the response rate of the faster firing baroreceptors in the model.The baseline heart rate ($h_0$) decreased slightly, to 88 bpm, from the initial 100 bpm.
+Within the estimated parameter set, the value of $N$, the baseline firing rate increased, and both $k_1$ and $\tau_1$, representing a change in the response rate of the faster firing baroreceptors in the model.
+Baroreceptors are generally divided into three groups, depending on their transmission speeds (on the order of half a second, five seonds, and five hundred seconds), but this model only includes two types.
+\cite{ottesen1996non} The increase in $\tau_1$ and the decrease in $\tau_2$ in the estimated parameter set may be the result of oversimplification.
+
+The baseline heart rate ($h_0$) decreased slightly, to 88 bpm, from the initial 100 bpm.
 Both $\tau_{ach}$ and $m_{ach}$, which determine how acetylcholine plays into the model, decreased dramatically in the estimated parameter set, with the reduction in $m_{ach}$ reducing the effects of a change in acetylcholine on heart rate, and the decrease in $\tau_{ach}$ increasing the rate of acetylcholine change.
@@
 As seen in Figure \ref{fig:paramscomparsionNLopt}, the estimated parameters slightly overestimate heart rate for this patient, but they perform far better than the original parameters, which severely underestimate the heart rate.
+This behavior is typical of the model on patients in this study.
 

### 2016-06-22 draft1.tex
added information about multiobjective parameters

@@
 The new parameter families reduced the averaged mean squared error even further than the parameters estimated using Nelder Mead-the ten best families for cluster 1 reduced the averaged mean squared error to 135 or less, and for cluster two, the averaged mean squared error was reduced to 178, or less.
+
+Both of the families had fairly similar values for $m_{nor}$ and $m_{ach}$, however, cluster 1 had a larger value of $N$, the resting firing rate than cluster 2.
+Cluster 1 contains the older patients, on average, suggesting that the resting firing rate may be age dependent.
+The differences in average heart rate between the clusters apparent in the clustering are appearent in the difference in $h_0$ between the two clusters, with cluster 2 having a faster base line heart rate, at $1.749 \pm .05$ beats per second, corresponding to $104.95 \pm 3$ beats per minute, compared to $74.4 \pm 4$ beats per minute for cluster 1.
 
@@
 \centering
+\scriptsize
 \caption{Original and estimated parameter values. The bolded parameters were held constant.}
@@
 Parameter Estimations from MIMIC II  Data & \textbf{1.5}      & 107.33 & \textbf{120} & 41.139 & \textbf{0.5}   & 14.062   & 202.925  & 2.94E-05     & 8.156        & \textbf{6}       & 1.474 & 0.165     & 5.20E-02  & \textbf{7}        \\ \hline
+\end{tabular}
+\end{table}
+
+\begin{table}[]
+\centering
+\scriptsize
+\caption{Average parameter values for the 10 best sets per cluster. Data is reported as mean $\pm$ stdev. Bolded columns were held constant.}
+\label{tab:MultiobjectiveParams}
+\begin{tabular}{|l|l|l|l|l|l|l|l|l|l|l|l|l|l|l|}
+\hline
+Cluster & \textbf{$\alpha$} & $N$          & \textbf{$M$} & $k_1$          & \textbf{$k_2$} & $\tau_1$       & $\tau_2$        & $\tau_{ach}$   & $\tau_{nor}$   & \textbf{$\beta$} & $h_0$           & $m_{nor}$      & $m_{ach}$       & \textbf{$\tau_d$} \\ \hline
+1       & \textbf{1.5}      & $50.36\pm 9$ & \textbf{120} & $.9566 \pm .3$ & \textbf{0.5}   & $.1854\pm .13$ & $127.31 \pm 46$ & $.5968 \pm .4$ & $.3176\pm .06$ & \textbf{6}       & $1.245\pm .006$ & $.1816\pm .02$ & $.0434\pm .004$ & \textbf{7}        \\ \hline
+2       & \textbf{1.5}      & $37.98\pm 9$ & \textbf{120} & $1.379\pm .7$  & \textbf{0.5}   & $.3621\pm .11$ & $82.88\pm 33$   & $.1410\pm .06$ & $.1688\pm.04$  & \textbf{6}       & $1.749\pm.05$   & $.1609\pm.02$  & $.0962\pm .06$  & \textbf{7}        \\ \hline
 \end{tabular}
@@
 \end{table} 
-\begin{table}[]
-\centering
-\caption{Sensitivity by clusters}
-\label{tab:clusterSensitivity}
-\begin{tabular}{|l|l|l|l|l|l|l|l|l|l|}
-\hline
-Cluster   & $\frac{dh}{dN}$ & $\frac{dh}{k_1}$ & $\frac{dh}{\tau_1}$ & $\frac{dh}{tau_2}$ & $\frac{dh}{\tau_{ach}}$ & $\frac{dh}{\tau_{nor}}$ & $\frac{dh}{h0}$ & $\frac{dh}{m_{nor}}$ & $\frac{dh}{m_{ach}}$ \\ \hline
-Overall   & 4.8352          & 0.2500           & 0.5088              & 0.0033             & 0.0204                  & 0.5720                  & 19.9294         & 1.1339               & 4.4651               \\ \hline
-cluster 1 & 4.7152          & 0.2456           & 0.5027              & 0.0033             & 0.0201                  & 0.5755                  & 19.4780         & 1.1438               & 4.3593               \\ \hline
-cluster 2 & 4.9137          & 0.2528           & 0.5128              & 0.0033             & 0.0206                  & 0.5697                  & 20.2248         & 1.1275               & 4.5344               \\ \hline
-\end{tabular}
-\end{table}
+%\begin{table}[]
+%\centering
+%\caption{Sensitivity by clusters}
+%\label{tab:clusterSensitivity}
+%\begin{tabular}{|l|l|l|l|l|l|l|l|l|l|}
+%\hline
+%Cluster   & $\frac{dh}{dN}$ & $\frac{dh}{k_1}$ & $\frac{dh}{\tau_1}$ & $\frac{dh}{tau_2}$ & $\frac{dh}{\tau_{ach}}$ & $\frac{dh}{\tau_{nor}}$ & $\frac{dh}{h0}$ & $\frac{dh}{m_{nor}}$ & $\frac{dh}{m_{ach}}$ \\ \hline
+%Overall   & 4.8352          & 0.2500           & 0.5088              & 0.0033             & 0.0204                  & 0.5720                  & 19.9294         & 1.1339               & 4.4651               \\ \hline
+%cluster 1 & 4.7152          & 0.2456           & 0.5027              & 0.0033             & 0.0201                  & 0.5755                  & 19.4780         & 1.1438               & 4.3593               \\ \hline
+%cluster 2 & 4.9137          & 0.2528           & 0.5128              & 0.0033             & 0.0206                  & 0.5697                  & 20.2248         & 1.1275               & 4.5344               \\ \hline
+%\end{tabular}
+%\end{table}
 \end{document}

### 2016-06-23 draft1.tex
changed single objective sensitivity table

@@
 where $n_s$ = 1, as $h$ is the only state variable and $n_t$ is the number of time points available for that patient.
+We calculated sensitivites of parameters that were optimized and not that of those held constant.
+\subsubsection*{Single Objective Sensitivity}
 From the original values provided by Olufsen and Ottesen, we found that $h_0$ was the most sensitive parameter, followed by $N$, $m_{ach}$, and $m_{nor}$ in that order.
@@
 The four parameters that were the most sensitive in our study were among the five most sensitive parameters found by Olufsen and Ottesen, as seen in \ref{tab:OSSCsoriginalParams}.
-With the new best parameters, the same four parameters are the most sensitive, as shown in \ref{tab:clusterSensitivity}. 
+\subsubsection*{Multiobjective Sensitivity} 
 For the multiobjective case, we used the same finite differences, but averaged them not only over the patients, but over the families of parameters-the ten best for each cluster.
@@
 \centering
-\caption{Average Overall State Sensitivity Coefficents Using Original Parameters}
+\caption{Average Overall State Sensitivity Coefficents For Single Objective Case}
 \label{tab:OSSCsoriginalParams}
-\begin{tabular}{lllllllll}
-$\frac{dh}{dN}$    & $\frac{dh}{dk_1}$  & $\frac{dh}{d\tau_1}$ & $\frac{dh}{d\tau_2}$ &$\frac{dh}{d\tau_{ach}}$ & $\frac{dh}{d\tau_{nor}}$ & $\frac{dh}{dh_0}$   & $\frac{dh}{dm_{nor}}$ & $\frac{dh}{dm_{ach}}$ \\
-15.3620 & 0.0792 & 0.0740  & 0.0156  & 0.0540   & 0.0006    & 19.9294 & 0.5740  & 14.5868
+\begin{tabular}{|l|l|l|l|l|l|l|l|l|l|}
+\hline
+Parameter set        & $S_{N}$    & $S_{k_1}$ & $S_{\tau_1}$ & $S_{\tau_2}$ & $S_{\tau_{ach}}$ & $S_{\tau_{nor}}$ & $S_{h_0}$   & $S_{m_{ach}}$ & $S_{m_{nor}}$ \\ \hline
+original parameters  &            &           &              &              &                  &                  &             &               &                \\ \hline
+estimated parameters & $1.75\pm4$ & $.76\pm3$ & $.87\pm3$    & $.39\pm2$    & $0.31\pm1.8$     & $4.18\pm1.5$     & $19.9\pm 6$ & $6.69\pm2$    & $.76\pm 1.8$   \\ \hline
 \end{tabular}

### 2016-06-23 draft1.tex
changed figures to pdfs and updated multiobjective sensitivity

@@
 For the multiobjective case, we used the same finite differences, but averaged them not only over the patients, but over the families of parameters-the ten best for each cluster.
-As with the single objective case, $h_0$ is the most sensitive parameter, followed by $N$, as seen in \ref{fig:MultiobjectiveSensitivityGraph}.
+As with the single objective case, $h_0$ is the most sensitive parameter, followed by $N$, as seen in \ref{fig:MultiobjectiveSensitivityGraph} for both clusters, but $N$ is about two times as sensitive in cluster 2 as in cluster one, whereass $h_0$ has nearly the same sensitivity for both clusters.
+If one were to rank the parameters by sensitivity, the ordering of the four most sensitive parameters between the two clusters is idenitical, however, $\tau_{nor}$ in cluster 2 is approximately an order of magnitude less sensitive than in cluster 1 and $\tau_{ach}$ is nearly three orders of magnitude less senstivie in cluster 2 compared to cluster 1.
+
+If we were to compare to the parameters estimated by Nelder Mead, we find similar sensitivites in $h_0$ and $N$ as well as in $m_{ach}$ and $m_{nor}$.
+The sensitivities in all of the $\tau$s included in the model is much smaller in the multiobjective case than in the single objective case.
+Part of this difference in sensitivity may arise from the differences in value found for these parameters by the different methods of optimization.
 \section*{Discussion}
@@
        \centering
-       \includegraphics[width=\linewidth]{figures/MultiobjectiveSenstivityGraph10familiesofNoErrorBars}
+       \includegraphics[width=\linewidth]{figures/MultiobjectiveSenstivityGraph10familiesofNoErrorBarsJun22}
        \caption{The black circles are from cluster 1, the grey x's represent cluster 2. Error bars are omitted for the sake of clarity.}

### 2016-06-24 draft1.tex
updated single objective sensitivity table and section

@@
 The four parameters that were the most sensitive in our study were among the five most sensitive parameters found by Olufsen and Ottesen, as seen in \ref{tab:OSSCsoriginalParams}.
+Between the estimated and the original parameters, the sensitivity of $h_0$ barely changed, but $N$ became much less sensitive in the estimated parameter set and $m_{ach}$ increased in sensitivity.
+This appearent change in sensitivity originates from the method of calculation, where the sum is divided by the value of the parameter.
+As $N$ increased in the estimated parameter set, dividing by this new value decreased the overal state sensitivity coefficent, while the opposite is true for $m_{ach}$.
 \subsubsection*{Multiobjective Sensitivity} 
@@
 \centering
+\scriptsize
 \caption{Average Overall State Sensitivity Coefficents For Single Objective Case}
@@
 Parameter set        & $S_{N}$     & $S_{k_1}$      & $S_{\tau_1}$  & $S_{\tau_2}$   & $S_{\tau_{ach}}$ & $S_{\tau_{nor}}$ & $S_{h_0}$   & $S_{m_{ach}}$ & $S_{m_{nor}}$ \\ \hline
-original parameters  &            &           &              &              &                  &                  &             &               &                \\ \hline
-estimated parameters & $1.75\pm4$ & $.76\pm3$ & $.87\pm3$    & $.39\pm2$    & $0.31\pm1.8$     & $4.18\pm1.5$     & $19.9\pm 6$ & $6.69\pm2$    & $.76\pm 1.8$   \\ \hline
+original parameters  & $15.26\pm4$ & $.0792\pm .04$ & $.0739\pm.04$ & $.0156\pm.004$ & $.0540\pm.013$   & $5.5E-4\pm4E-4$ & $19.9\pm 6$ & $.574\pm.7$   & $14.6\pm4$    \\ \hline
+estimated parameters & $2.09\pm5$  & $.83\pm3$      & $1.11\pm3$    & $.48\pm2$      & $0.43\pm2$       & $4.18\pm1.6$     & $19.8\pm 6$ & $6.69\pm2$    & $.71\pm 1.6$  \\ \hline
 \end{tabular}

### 2016-06-24 draft1.tex
added some new information to intro

@@
 The rate at which a heart beats is determined, in part, by the sympathetic and parasympathetic portions of the nervous system.
-When the sympathetic nervous system is stimulated, it releases epinephrine and norepinephrine  which increase heart rate.
-The parasympathetic system releases acetylcholine which decreases heart rate.
+When the sympathetic nervous system is stimulated, it releases epinephrine and norepinephrine.
+Norepinephrine increases heart rate by increasing the rate of depolarization in pacemaker cells and increases contractability through the phosphorlation of several types of receptors.\cite{boron2016medical} The parasympathetic system releases acetylcholine which decreases heart rate and contractibility through its binding to M$_2$ receptors.
 These two systems, often described as an accelerator and a brake, are not totally independent on each other, rather, they interact through second messengers cAMP and cGMP.
@@
 \cite{ottesen1997modelling}
+
+This control process leads to a negative feedback loop, where if the mean arterial pressure increases, vasodilation and bradycardia occur, whereas if the mean arterial pressure decreases, the opposite, vasoconstriction and tachycardia occur.
+\cite{boron2016medical}
 

### 2016-06-27 draft1.tex
Updated single objective sensitivity to include all patients

@@
 Parameter set        & $S_{N}$     & $S_{k_1}$      & $S_{\tau_1}$  & $S_{\tau_2}$   & $S_{\tau_{ach}}$ & $S_{\tau_{nor}}$ & $S_{h_0}$   & $S_{m_{ach}}$ & $S_{m_{nor}}$ \\ \hline
-original parameters  & $15.26\pm4$ & $.0792\pm .04$ & $.0739\pm.04$ & $.0156\pm.004$ & $.0540\pm.013$   & $5.5E-4\pm4E-4$ & $19.9\pm 6$ & $.574\pm.7$   & $14.6\pm4$    \\ \hline
-estimated parameters & $2.09\pm5$  & $.83\pm3$      & $1.11\pm3$    & $.48\pm2$      & $0.43\pm2$       & $4.18\pm1.6$     & $19.8\pm 6$ & $6.69\pm2$    & $.71\pm 1.6$  \\ \hline
+original parameters  & $15.26\pm4$ & $.0792\pm .04$ & $.0739\pm.04$ & $.0156\pm.004$ & $.0540\pm.013$   & $0.00055\pm4E-4$ & $19.9\pm 6$ & $.574\pm.7$   & $14.6\pm4$    \\ \hline
+estimated parameters & $2.24\pm5$  & $.89\pm3$      & $1.20\pm4$    & $.61\pm3$      & $0.52\pm2$       & $4.34\pm1.9$     & $20.1\pm 6$ & $6.79\pm2$    & $.83\pm 1.9$  \\ \hline
 \end{tabular}

### 2016-06-29 draft1.tex
updated OSSC table to include parameters held constant

--- 2016-06-24 draft1.tex
@@
 
+%\begin{table}[h]
+%\centering
+%\scriptsize
+%\caption{Average Overall State Sensitivity Coefficents For Single Objective Case}
+%\label{tab:OSSCsoriginalParams}
+%\begin{tabular}{|l|l|l|l|l|l|l|l|l|l|}
+%\hline
+%Parameter set        & $S_{N}$     & $S_{k_1}$      & $S_{\tau_1}$  & $S_{\tau_2}$   & $S_{\tau_{ach}}$ & $S_{\tau_{nor}}$ & $S_{h_0}$   & $S_{m_{ach}}$ & $S_{m_{nor}}$ \\ \hline
+%original parameters  & $15.26\pm4$ & $.0792\pm .04$ & $.0739\pm.04$ & $.0156\pm.004$ & $.0540\pm.013$   & $0.00055\pm4E-4$ & $19.9\pm 6$ & $.574\pm.7$   & $14.6\pm4$    \\ \hline
+%estimated parameters & $2.24\pm5$  & $.89\pm3$      & $1.20\pm4$    & $.61\pm3$      & $0.52\pm2$       & $4.34\pm1.9$     & $20.1\pm 6$ & $6.79\pm2$    & $.83\pm 1.9$  \\ \hline
+%\end{tabular}
+%\end{table}
+
 \begin{table}[h]
 \centering
-\scriptsize
+\tiny
 \caption{Average Overall State Sensitivity Coefficents For Single Objective Case}
 \label{tab:OSSCsoriginalParams}
-\begin{tabular}{|l|l|l|l|l|l|l|l|l|l|}
+\begin{tabular}{|l|l|l|l|l|l|l|l|l|l|l|l|l|l|l|}
 \hline
-Parameter set        & $S_{N}$     & $S_{k_1}$      & $S_{\tau_1}$  & $S_{\tau_2}$   & $S_{\tau_{ach}}$ & $S_{\tau_{nor}}$ & $S_{h_0}$   & $S_{m_{ach}}$ & $S_{m_{nor}}$ \\ \hline
-original parameters  & $15.26\pm4$ & $.0792\pm .04$ & $.0739\pm.04$ & $.0156\pm.004$ & $.0540\pm.013$   & $5.5E-4\pm4E-4$ & $19.9\pm 6$ & $.574\pm.7$   & $14.6\pm4$    \\ \hline
-estimated parameters & $2.09\pm5$  & $.83\pm3$      & $1.11\pm3$    & $.48\pm2$      & $0.43\pm2$       & $4.18\pm1.6$     & $19.8\pm 6$ & $6.69\pm2$    & $.71\pm 1.6$  \\ \hline
+Parameter set        & $S_{\alpha}$ & $S_{N}$     & $S_{M}$        & $S_{k_1}$      & $S_{k_2}$      & $S_{\tau_1}$  & $S_{\tau_2}$   & $S_{\tau_{ach}}$ & $S_{\tau_{nor}}$ & $S_{\beta}$  & $S_{h_0}$   & $S_{m_{ach}}$ & $S_{m_{nor}}$ & $S_{\tau_d}$     \\ \hline
+original parameters  & $3.99\pm 2$  & $15.26\pm4$ & $.0069\pm.002$ & $.0792\pm .04$ & $142.61\pm 30$ & $.0739\pm.04$ & $.0156\pm.004$ & $.0540\pm.013$   & $0.00055\pm4E-4$ & $.113\pm.13$ & $19.9\pm 6$ & $.574\pm.7$   & $14.6\pm4$    & $.085\pm.09$     \\ \hline
+estimated parameters & $70\pm200$   & $2.24\pm5$  & $.322\pm.6$    & $.89\pm3$      & $108\pm400$    & $1.20\pm4$    & $.61\pm3$      & $0.52\pm2$       & $4.34\pm1.9$     & $.234\pm.19$ & $20.1\pm 6$ & $6.79\pm2$    & $.83\pm 1.9$  & $2.08E-7\pm2E-7$ \\ \hline
 \end{tabular}

### 2016-06-29 draft1.tex
updated single objective OSSC analysis

@@
 This control process leads to a negative feedback loop, where if the mean arterial pressure increases, vasodilation and bradycardia occur, whereas if the mean arterial pressure decreases, the opposite, vasoconstriction and tachycardia occur.
-\cite{boron2016medical}
+\cite{boron2016medical} In patients suffering from hypertension, the controlers in this loop get reset to the higher pressure, but the set point can be changed again where blood pressure returns to normal levels.
+\cite{oparil2003pathogenesis}
 
@@
 
-$k_1$ and $\beta$ were held constant as to reduce the search space.
+$k_2$ and $\beta$ were held constant as to reduce the search space, as done by Olufsen and Ottesen.
 $\alpha$, the parameter used to smooth the pressure data was held constant at 1.5 to smooth the data with a minimal phase shift.
@@
 \subsubsection*{Single Objective Sensitivity}
-From the original values provided by Olufsen and Ottesen, we found that $h_0$ was the most sensitive parameter, followed by $N$, $m_{ach}$, and $m_{nor}$ in that order.
+From the original values provided by Olufsen and Ottesen, we found that $h_0$ was the most sensitive parameter, followed by $N$, $m_{ach}$, and $m_{nor}$ in that order, of the parameters estimated.
 Of the four most sensitive equations, three of them ($h_0$, $m_{ach}$ and $m_{nor}$) appear directly in the equation for heart rate, and $N$ indirectly appears, as both $c_{ach}$ and $c_{nor}$ are functions of $N$.
-The four parameters that were the most sensitive in our study were among the five most sensitive parameters found by Olufsen and Ottesen, as seen in \ref{tab:OSSCsoriginalParams}.
+The four estimated parameters that were the most sensitive in our study were among the five most sensitive parameters found by Olufsen and Ottesen, as seen in \ref{tab:OSSCsoriginalParams}.
+However, $k_2$ showed a very high degree of sensitivity in both the original and estimated parameter set, even more so than any of the estimated parameters.
+It also had the highest varriance in sensitivity, due to the fact that some patients exhibited a sensitivity to $k_2$ that was an order of magnitude larger than the mean.
+
 Between the estimated and the original parameters, the sensitivity of $h_0$ barely changed, but $N$ became much less sensitive in the estimated parameter set and $m_{ach}$ increased in sensitivity.

### 2016-06-29 draft1.tex
started discussion section

@@
 \section*{Discussion}
-The Discussion should be succinct and must not contain subheadings.
+We estimated three sets of parameters that can be used to predict heart rates of patients in intensive care units.
+The accuracy of the model was much improved when patients were grouped by their characterstics.
+
+The model used has several shortcommings.
+This model oversimplifies the process of heart rate regulation.
+It does not include cardiac sypathetic afferrent reflex, which is sensitive to both mechanical and chemical inputs.
+\cite{du2007love} It can only be used for short time scales, as it does not include the slower acting renin-agiotensin-aldostrone system, which determines blood volume, and therefore blood pressure.
+\cite{boron2016medical} Nevertheless, it does captures the dynamics of the system at short time scales.
 \section*{Methods}

### 2016-06-30 draft1.tex
Updated sensitivity table to include results from all patients

@@
 Differential equations were solved using the ODE package, with solvers ode23 and ode78, with an absolute and relative tolerance of $10^{-8}$.
+
+Since pressure data was given in a discrete format, we linearly interpolated between values when neccessary, as pressure is an input to the system of differential equations.
 \bibliography{draft1}
@@
 original parameters  & $3.99\pm 2$  & $15.26\pm4$ & $.0069\pm.002$ & $.0792\pm .04$ & $142.61\pm 30$ & $.0739\pm.04$ & $.0156\pm.004$ & $.0540\pm.013$   & $0.00055\pm4E-4$ & $.113\pm.13$ & $19.9\pm 6$ & $.574\pm.7$   & $14.6\pm4$    & $.085\pm.09$     \\ \hline
-estimated parameters & $70\pm200$   & $2.24\pm5$  & $.322\pm.6$    & $.89\pm3$      & $108\pm400$    & $1.20\pm4$    & $.61\pm3$      & $0.52\pm2$       & $4.34\pm1.9$     & $.234\pm.19$ & $20.1\pm 6$ & $6.79\pm2$    & $.83\pm 1.9$  & $2.08E-7\pm2E-7$ \\ \hline
+estimated parameters & $78\pm250$   & $2.24\pm5$  & $.329\pm.6$    & $.89\pm3$      & $86\pm400$     & $1.20\pm4$    & $.61\pm3$      & $0.52\pm2$       & $4.34\pm1.9$     & $.52\pm5$    & $20.1\pm 6$ & $6.79\pm2$    & $.83\pm 1.9$  & $4.13E-7\pm3E-6$ \\ \hline
 \end{tabular}

### 2016-07-06 draft1.tex
updated multiobjective sensitivity graph to include extended parameters

@@
        \centering
-       \includegraphics[width=\linewidth]{figures/MultiobjectiveSenstivityGraph10familiesofNoErrorBarsJun22}
-       \caption{The black circles are from cluster 1, the grey x's represent cluster 2. Error bars are omitted for the sake of clarity.}
+       \includegraphics[width=\linewidth]{figures/MultiobjectiveSenstivityGraph10familiesofNoErrorBarsWithExtendedSensivities}
+       \caption{The black marks are from cluster 1, the grey markss represent cluster 2.
+Error bars are omitted for the sake of clarity.
+Parameters represented by an x were allowed to vary, and parameters represented by a star were held constant.}
         \label{fig:MultiobjectiveSensitivityGraph}

### 2016-07-07 draft1.tex
Included cluster 2 extended sensitivity.

@@
 If one were to rank the parameters by sensitivity, the ordering of the four most sensitive parameters between the two clusters is idenitical, however, $\tau_{nor}$ in cluster 2 is approximately an order of magnitude less sensitive than in cluster 1 and $\tau_{ach}$ is nearly three orders of magnitude less senstivie in cluster 2 compared to cluster 1.
+Of the parameters held constant during optimization, $k_2$ showed the highest degree of sensitivity in both clusters, which corresponds to the sensitivity found in the single objective case.
 
@@
 Error bars are omitted for the sake of clarity.
-Parameters represented by an x were allowed to vary, and parameters represented by a star were held constant.}
+Parameters represented by an square were allowed to vary, and parameters represented by a diamond were held constant.}
         \label{fig:MultiobjectiveSensitivityGraph}

### 2016-07-14 draft1.tex
added some to discussion

@@
 The accuracy of the model was much improved when patients were grouped by their characterstics.
+This improvement in accuracy is likely mostly based on that the resting heart rate of the patients in clusters more accurately matched $h_0$, the resting heart rate in the model, since $h_0$ was shown to be the most sensitive parameter in the model.
+
+We discovered that one of the parameters found to be fairly unsensitive in previous work, $k_2$, was surprisingly sensitive in this work.
+It was held constant during optimization as previous work pointed to it being a very unsensitive parameter, but it proved to be very sensitive in all cases.
+Future work may involve estimating a better value for the gain on this type of barroreceptor.
 
@@
 Differential equations were solved using the ODE package, with solvers ode23 and ode78, with an absolute and relative tolerance of $10^{-8}$.
+Single objective optimization was perfomed using the Nelder Mead algorithm, as implimented in the Julia package NLopt, and multiobjective optimization was performed using the POETs package.
 
 Since pressure data was given in a discrete format, we linearly interpolated between values when neccessary, as pressure is an input to the system of differential equations.
+We used the pressure data recorded at 1 Hz as well as the heart rate data at 1 Hz from the MIMIC II Waveform Database, numerics section.
+We extracted data from the database using rdsamp and stored it locally as text files that were then read into the model.
 \bibliography{draft1}
@@
        \includegraphics[width=\linewidth]{figures/MultiobjectiveSenstivityGraph10familiesofNoErrorBarsWithExtendedSensivities}
-       \caption{The black marks are from cluster 1, the grey markss represent cluster 2.
+       \caption{The black marks are from cluster 1, the grey marks represent cluster 2.
 Error bars are omitted for the sake of clarity.

### 2016-08-30 draft1.tex
Added model formulation

@@
 Additionally, many previous studies were performed with healthy patients, in contrast, this study focuses on intensive care unit patients.
+\subsection{Model Formulation}
+The model assumes changes in arterial wall stretch are proportionate to filtered blood pressure, $\dot{\bar{p}} = \alpha(-\bar p + p)$
+where $\alpha$ is a gain, $\bar p$ is the filtered blood pressure, and $p$ is the patient's measured blood pressure.
+From $\bar p$, we can predict the nervous system firing rate, $n = \sum_i n_i + N,~i = 1,2$
+where N is a baseline firing rate and $n_i$ corresponds to the firing rate of nerve fibers of type $i$.
+The firing rates for nerves of type $i$ are computed as:
+\begin{equation}
+\frac{dn_i}{dt} = \kappa_i \frac{d \bar p}{dt} \frac{n(M-n)}{(M/2)^2}-\frac{n_i}{\tau_i}, i = 1,2
+\end{equation}
+where $M$ is the maximum firing rate, and $\tau_i$ is the time scale for nerves of type $i$.
+The firing rate information is compiled by the central nervous system, which then determines the the sympathetic and parasympathetic outputs, $f_{sym}$ and $f_{par}$, respectively.
+The parasympathetic output is given by $f_{par} = n/M$, while sympathetic output takes the form:
+\begin{equation}
+f_{sym} = \frac{1-n(t-\tau_d)/M}{1+\beta f_{par}}
+\end{equation}
+With these outputs, we can determine the dimensionless concentrations of acetylcholine, $c_{ach}$ and noradrenaline, $c_{nor}$:
+\begin{equation}
+\label{dcnordt}
+\frac{dc_{j}}{dt} = \frac{f_{i}-c_{j}}{\tau_{j}}
+\end{equation}
+where each $\tau_{j}$ represents a time scale.
+Finally, the heart rate is calculated from $h_0$, the intrinsic heart rate, and $m_{nor}$ and $m_{ach}$,
+which are weights for the contributions of acetylcholine and norepinephrine to heart rate $h = h_0(1+m_{nor}c_{nor} - m_{ach}c_{ach})$.
 \section*{Results}
@@
 
-The model used has several shortcommings.
+The model used has several shortcomings.
 This model oversimplifies the process of heart rate regulation.
@@
 \cite{du2007love} It can only be used for short time scales, as it does not include the slower acting renin-agiotensin-aldostrone system, which determines blood volume, and therefore blood pressure.
-\cite{boron2016medical} Nevertheless, it does captures the dynamics of the system at short time scales.
+\cite{boron2016medical} The model only considers two vassoctive agents when there are many more and does not attempt to capture the effects of respiration rate.
+Nevertheless, it does captures the dynamics of the system at short time scales.
 \section*{Methods}

### 2016-08-31 draft1.tex
Added results about allowing k2 and beta to vary during optimization

@@
 $k_2$ and $\beta$ were held constant as to reduce the search space, as done by Olufsen and Ottesen.
+Additionally, when they were allowed to vary during the optimization, the parameter set generated had very small values for $m_{ach}$ and $m__{nor}$, resulting in a predicted heart rate that was basically constant at $h_0$.
+While this set of the parameters did mimimize the MSE, it did not capture the intrinsic varation in heart rate.
 $\alpha$, the parameter used to smooth the pressure data was held constant at 1.5 to smooth the data with a minimal phase shift.

### 2016-09-14 draft1.tex
added section on global sensitivity analysis

--- 2016-08-30 draft1.tex
@@
 $k_2$ and $\beta$ were held constant as to reduce the search space, as done by Olufsen and Ottesen.
+Additionally, when they were allowed to vary during the optimization, the parameter set generated had very small values for $m_{ach}$ and $m_{nor}$, resulting in a predicted heart rate that was basically constant at $h_0$.
+While this set of the parameters did mimimize the MSE, it did not capture the intrinsic varation in heart rate.
 $\alpha$, the parameter used to smooth the pressure data was held constant at 1.5 to smooth the data with a minimal phase shift.
@@
 
-\subsection*{Sensitivity Using Finite Differences}
+\subsection*{Local Sensitivity Using Finite Differences}
 The derivatives of all parameters were estimated using central differences. 
@@
 Part of this difference in sensitivity may arise from the differences in value found for these parameters by the different methods of optimization.
+
+\subsection{Global Sensitivity}
+We used the Morris method of sampling to generate a family of parameter sets, ranging from $\pm$25\% of the nominal parameter values, with the exception of $h_0$, which was bounded between [.8, 2.0], corresponding to a resting heart rate between 48 and 120 bpm.\cite{morris1991factorial} We found that this range of parameters resulted in physiologically acceptable heart rates.
+As heart rates are a function of time, it was neccessary for us to collapse the time dimension to perform sensitivity analysis.
+We used fPCA (functional principal component analysis) to convert the heart rate curves into weights ($\omega_i$) and eigenfunctions $(\xi_i(t))$ so that $h_j(t)=\sum_i \omega_i \xi_i(t)$, and then performed sensitivity analysis on the weights.
+\cite{sumner2012methodology}
+
 \section*{Discussion}

### 2016-09-20 draft1.tex
added Morris plots

@@
 As heart rates are a function of time, it was neccessary for us to collapse the time dimension to perform sensitivity analysis.
-We used fPCA (functional principal component analysis) to convert the heart rate curves into weights ($\omega_i$) and eigenfunctions $(\xi_i(t))$ so that $h_j(t)=\sum_i \omega_i \xi_i(t)$, and then performed sensitivity analysis on the weights.
-\cite{sumner2012methodology}
+We used fPCA (functional principal component analysis) to convert the heart rate curves into weights ($\omega_i$) and eigenfunctions $(\xi_i(t))$ so that $h_j(t)=\sum_i \omega_i \xi_i(t)$, and then performed sensitivity analysis on the weights generated for each patient.
+\cite{sumner2012methodology} We then grouped the patients by cluster, and calculated the average of $\mu^*$, the estimate of the mean of distrubution of the absolute values of the elementry effects.
+For each cluster, we calculated the overall standard deviation, $\sigma$, by summing all of the variances and then taking the square root.
+\cite{1_normal sum distribution} The results are shown in \ref{fig:Cluster1Morris} and \ref{fig:Cluster2Morris}.
 
@@
 Single objective optimization was perfomed using the Nelder Mead algorithm, as implimented in the Julia package NLopt, and multiobjective optimization was performed using the POETs package.
+FPCA calculations were carried out through the use of the Rcall package using the R package FDApace.
 
@@
 
+\begin{figure}[ht]
+                \centering
+        \includegraphics[width=\linewidth]{figures/MorrisPlotN100Cluster1}
+       \caption{Cluster 1 Morris Plot, with N =100}
+       \label{fig:Cluster1Morris}
+\end{figure}
+
+\begin{figure}[ht]
+                \centering
+        \includegraphics[width=\linewidth]{figures/MorrisPlotN100Cluster2}
+       \caption{Cluster 2 Morris Plot, with N =100}
+       \label{fig:Cluster2Morris}
+\end{figure}
+
 %\begin{table}[ht]

### 2016-09-20 draft1.tex
a few changes to sensitivity section

@@
 \cite{1_normal sum distribution} The results are shown in \ref{fig:Cluster1Morris} and \ref{fig:Cluster2Morris}.
+The results are nearly identical between the two clusters, with the ranking of the six most sensitive paramaters (as given by the largest values of $\mu^*$) being the same in both clusters, with $m_{nor}$ being the parameter with the largest effect on the model's output.
+The parameter with the largest standard deviation, indicative of it depending on other paramters in the model, was the resting heart rate.
 

### 2016-09-26 draft1.tex
changed global sensitivty results to use the N=20 case

--- 2016-09-20 draft1.tex
@@
-              
 \documentclass[fleqn,10pt]{wlscirep}
@@
 \author[1,*]{Jeffrey D. Varner}
-\affil[1]{Affiliation, School of Chemical and Biomolecular Engineering, Ithaca NY, 14852, USA}
+\affil[1]{School of Chemical and Biomolecular Engineering, Ithaca NY, 14852, USA}
 \affil[*]{jdv27@cornell.edu}
@@
 \cite{ottesen1997modelling}
-
 This control process leads to a negative feedback loop, where if the mean arterial pressure increases, vasodilation and bradycardia occur, whereas if the mean arterial pressure decreases, the opposite, vasoconstriction and tachycardia occur.
@@
 \cite{oparil2003pathogenesis}
-
 Olufsen and Ottesen have developed models of heart rate based on blood pressure measurements.\cite{olufsen2013practical} From the blood pressure measurement, the model predicts a firing rate for different types of receptors, which is then used to predict the response times of the parasympathetic and sympathetic nervous system.
 The response times are then used to predict dimensionless norepinephrine and acetylcholine concentrations, which are finally used to predict heart rate.
-
 We used the MIMIC II Waveform Database as the source of the pressure and heart rate data used to test this model.\cite{saeed2011multiparameter} MIMIC contains de-identified data from patients who visited the Beth Israel Deaconess Medical Center ICU.
@@
 We minimized averaged the mean squared error per patient using the Nelder Mead algorithm as implemented in Julia package NLopt.
-We forced all of the parameters to be non-negative as to be biologically correct.The original and estimated values of the parameters are shown in Table \ref{tab:SingleParamsNLopt}.
+We forced all of the parameters to be non-negative as to be biologically correct.
+The original and estimated values of the parameters are shown in Table \ref{tab:SingleParamsNLopt}.
 %The value of $\tau_2$ decreased dramatically in the estimated parameter set, increasing the effect of $n_1$ on the firing rate.
@@
 The change in parameters reduced the mean squared error, averaged over all of the selected patients, from 1167 to 349.
-
 Within the estimated parameter set, the value of $N$, the baseline firing rate increased, and both $k_1$ and $\tau_1$, representing a change in the response rate of the faster firing baroreceptors in the model.
-Baroreceptors are generally divided into three groups, depending on their transmission speeds (on the order of half a second, five seonds, and five hundred seconds), but this model only includes two types.
+Baroreceptors are generally divided into three groups, depending on their transmission speeds (on the order of half a second, five seconds, and five hundred seconds), but this model only includes two types.
 \cite{ottesen1996non} The increase in $\tau_1$ and the decrease in $\tau_2$ in the estimated parameter set may be the result of oversimplification.
-
 The baseline heart rate ($h_0$) decreased slightly, to 88 bpm, from the initial 100 bpm.
@@
 This behavior is typical of the model on patients in this study.
-
 $k_2$ and $\beta$ were held constant as to reduce the search space, as done by Olufsen and Ottesen.
 Additionally, when they were allowed to vary during the optimization, the parameter set generated had very small values for $m_{ach}$ and $m_{nor}$, resulting in a predicted heart rate that was basically constant at $h_0$.
-While this set of the parameters did mimimize the MSE, it did not capture the intrinsic varation in heart rate.
+While this set of the parameters did minimize the MSE, it did not capture the intrinsic variation in heart rate.
 $\alpha$, the parameter used to smooth the pressure data was held constant at 1.5 to smooth the data with a minimal phase shift.
@@
 \cite{baker1997pitfalls}
-
 \subsection*{Clustering}
@@
 \subsection*{Multidimensional Optimization}
-We used the Julia language POETs package, which combines simulated anealing with Pareto optimality to generate families of best parameters.\cite{bassen2016jupoets} Using the two clusters formed by k-means, we minimized the averaged mean squared error to create these parameter families.
+We used the Julia language POETs package, which combines simulated annealing with Pareto optimality to generate families of best parameters.\cite{bassen2016jupoets} Using the two clusters formed by k-means, we minimized the averaged mean squared error to create these parameter families.
 We utilized a monotonically decreasing cooling schedule with five iterations at each temperature.
@@
 The new parameter families reduced the averaged mean squared error even further than the parameters estimated using Nelder Mead-the ten best families for cluster 1 reduced the averaged mean squared error to 135 or less, and for cluster two, the averaged mean squared error was reduced to 178, or less.
-
 Both of the families had fairly similar values for $m_{nor}$ and $m_{ach}$, however, cluster 1 had a larger value of $N$, the resting firing rate than cluster 2.
 Cluster 1 contains the older patients, on average, suggesting that the resting firing rate may be age dependent.
-The differences in average heart rate between the clusters apparent in the clustering are appearent in the difference in $h_0$ between the two clusters, with cluster 2 having a faster base line heart rate, at $1.749 \pm .05$ beats per second, corresponding to $104.95 \pm 3$ beats per minute, compared to $74.4 \pm 4$ beats per minute for cluster 1.
+The differences in average heart rate between the clusters apparent in the clustering are apparent in the difference in $h_0$ between the two clusters, with cluster 2 having a faster base line heart rate, at $1.749 \pm .05$ beats per second, corresponding to $104.95 \pm 3$ beats per minute, compared to $74.4 \pm 4$ beats per minute for cluster 1.
 
@@
 where $n_s$ = 1, as $h$ is the only state variable and $n_t$ is the number of time points available for that patient.
-We calculated sensitivites of parameters that were optimized and not that of those held constant.
+We calculated sensitivities of parameters that were optimized and not that of those held constant.
 \subsubsection*{Single Objective Sensitivity}
@@
 However, $k_2$ showed a very high degree of sensitivity in both the original and estimated parameter set, even more so than any of the estimated parameters.
-It also had the highest varriance in sensitivity, due to the fact that some patients exhibited a sensitivity to $k_2$ that was an order of magnitude larger than the mean.
-
+It also had the highest variance in sensitivity, due to the fact that some patients exhibited a sensitivity to $k_2$ that was an order of magnitude larger than the mean.
 Between the estimated and the original parameters, the sensitivity of $h_0$ barely changed, but $N$ became much less sensitive in the estimated parameter set and $m_{ach}$ increased in sensitivity.
-This appearent change in sensitivity originates from the method of calculation, where the sum is divided by the value of the parameter.
-As $N$ increased in the estimated parameter set, dividing by this new value decreased the overal state sensitivity coefficent, while the opposite is true for $m_{ach}$.
+This apparent change in sensitivity originates from the method of calculation, where the sum is divided by the value of the parameter.
+As $N$ increased in the estimated parameter set, dividing by this new value decreased the overall state sensitivity coefficient, while the opposite is true for $m_{ach}$.
 \subsubsection*{Multiobjective Sensitivity} 
@@
 As with the single objective case, $h_0$ is the most sensitive parameter, followed by $N$, as seen in \ref{fig:MultiobjectiveSensitivityGraph} for both clusters, but $N$ is about two times as sensitive in cluster 2 as in cluster one, whereass $h_0$ has nearly the same sensitivity for both clusters.
-If one were to rank the parameters by sensitivity, the ordering of the four most sensitive parameters between the two clusters is idenitical, however, $\tau_{nor}$ in cluster 2 is approximately an order of magnitude less sensitive than in cluster 1 and $\tau_{ach}$ is nearly three orders of magnitude less senstivie in cluster 2 compared to cluster 1.
+If one were to rank the parameters by sensitivity, the ordering of the four most sensitive parameters between the two clusters is identical, however, $\tau_{nor}$ in cluster 2 is approximately an order of magnitude less sensitive than in cluster 1 and $\tau_{ach}$ is nearly three orders of magnitude less sensitive in cluster 2 compared to cluster 1.
 Of the parameters held constant during optimization, $k_2$ showed the highest degree of sensitivity in both clusters, which corresponds to the sensitivity found in the single objective case.
-
-If we were to compare to the parameters estimated by Nelder Mead, we find similar sensitivites in $h_0$ and $N$ as well as in $m_{ach}$ and $m_{nor}$.
+If we were to compare to the parameters estimated by Nelder Mead, we find similar sensitivities in $h_0$ and $N$ as well as in $m_{ach}$ and $m_{nor}$.
 The sensitivities in all of the $\tau$s included in the model is much smaller in the multiobjective case than in the single objective case.
 Part of this difference in sensitivity may arise from the differences in value found for these parameters by the different methods of optimization.
-
 \subsection{Global Sensitivity}
 We used the Morris method of sampling to generate a family of parameter sets, ranging from $\pm$25\% of the nominal parameter values, with the exception of $h_0$, which was bounded between [.8, 2.0], corresponding to a resting heart rate between 48 and 120 bpm.\cite{morris1991factorial} We found that this range of parameters resulted in physiologically acceptable heart rates.
-As heart rates are a function of time, it was neccessary for us to collapse the time dimension to perform sensitivity analysis.
+As heart rates are a function of time, it was necessary for us to collapse the time dimension to perform sensitivity analysis.
 We used fPCA (functional principal component analysis) to convert the heart rate curves into weights ($\omega_i$) and eigenfunctions $(\xi_i(t))$ so that $h_j(t)=\sum_i \omega_i \xi_i(t)$, and then performed sensitivity analysis on the weights generated for each patient.
-\cite{sumner2012methodology} We then grouped the patients by cluster, and calculated the average of $\mu^*$, the estimate of the mean of distrubution of the absolute values of the elementry effects.
+\cite{sumner2012methodology} We then grouped the patients by cluster, and calculated the average of $\mu^*$, the estimate of the mean of distribution of the absolute values of the elementary effects.
 For each cluster, we calculated the overall standard deviation, $\sigma$, by summing all of the variances and then taking the square root.
 \cite{1_normal sum distribution} The results are shown in \ref{fig:Cluster1Morris} and \ref{fig:Cluster2Morris}.
-
+The results are nearly identical between the two clusters, with the ranking of the six most sensitive parameters (as given by the largest values of $\mu^*$) being the same in both clusters, with $h_0$ being the parameter with the largest effect on the model's output.
+It was also the parameter with the largest standard deviation, indicitive of it depending on the other parameters in the model.
+This agrees with the results of the local sensitivity analysis, where $h_0$ was the parameter with the largest impact on the model's output.
+However, two parameters, ($N$ and $k_2$), which showed high sensitivity in the local case, did not in the global case, but $m_{ach}$, which was sensitive in the local case was also found to be a strong determinant of output in the global case.
+%, with $m_{nor}$ being the parameter with the largest effect on the model's output.
+The parameter with the largest standard deviation, indicative of it depending on other parameters in the model, was the resting heart rate.
+It is somewhat surprising that $m_{nor}$ is the most sensitive parameter, as it was not the most sensitive in the local sensitivity analyses, but it was among four most sensitive in both the local sensitivity analyses for the clusters, as well the original parameter set.
 \section*{Discussion}
 We estimated three sets of parameters that can be used to predict heart rates of patients in intensive care units.
-The accuracy of the model was much improved when patients were grouped by their characterstics.
+The accuracy of the model was much improved when patients were grouped by their characteristics.
 This improvement in accuracy is likely mostly based on that the resting heart rate of the patients in clusters more accurately matched $h_0$, the resting heart rate in the model, since $h_0$ was shown to be the most sensitive parameter in the model.
-
-We discovered that one of the parameters found to be fairly unsensitive in previous work, $k_2$, was surprisingly sensitive in this work.
-It was held constant during optimization as previous work pointed to it being a very unsensitive parameter, but it proved to be very sensitive in all cases.
+We discovered that one of the parameters found to be fairly insensitive in previous work, $k_2$, was surprisingly sensitive in this work.
+It was held constant during optimization as previous work pointed to it being a very insensitive parameter, but it proved to be very sensitive in all cases.
 Future work may involve estimating a better value for the gain on this type of barroreceptor.
-
 The model used has several shortcomings.
 This model oversimplifies the process of heart rate regulation.
-It does not include cardiac sypathetic afferrent reflex, which is sensitive to both mechanical and chemical inputs.
+It does not include cardiac sympathetic afferent reflex, which is sensitive to both mechanical and chemical inputs.
 \cite{du2007love} It can only be used for short time scales, as it does not include the slower acting renin-agiotensin-aldostrone system, which determines blood volume, and therefore blood pressure.
@@
 Differential equations were solved using the ODE package, with solvers ode23 and ode78, with an absolute and relative tolerance of $10^{-8}$.
-Single objective optimization was perfomed using the Nelder Mead algorithm, as implimented in the Julia package NLopt, and multiobjective optimization was performed using the POETs package.
+Single objective optimization was performed using the Nelder Mead algorithm, as implimented in the Julia package NLopt, and multiobjective optimization was performed using the POETs package.
 FPCA calculations were carried out through the use of the Rcall package using the R package FDApace.
-
-Since pressure data was given in a discrete format, we linearly interpolated between values when neccessary, as pressure is an input to the system of differential equations.
+Since pressure data was given in a discrete format, we linearly interpolated between values when necessary, as pressure is an input to the system of differential equations.
 We used the pressure data recorded at 1 Hz as well as the heart rate data at 1 Hz from the MIMIC II Waveform Database, numerics section.
@@
 \section*{Author contributions statement}
-Must include all authors, identified by initials, for example:
-A.A. conceived the experiment(s),  A.A. and B.A. conducted the experiment(s), C.A. and D.A. analysed the results.  All authors reviewed the manuscript. 
+R.L. conducted the experiments, R.L. and J.V. analysed the results.  All authors reviewed the manuscript. 
 \section*{Additional information}
@@
 This statement must be included in the submitted article file.
-
 %\begin{figure}[ht]
@@
 %\end{figure}
-
 \begin{figure}[ht]
@@
 \end{figure}
-
 \begin{figure}[ht]
                 \centering
-        \includegraphics[width=\linewidth]{figures/MorrisPlotN100Cluster1}
-       \caption{Cluster 1 Morris Plot, with N =100}
+        \includegraphics[width=\linewidth]{figures/MorrisPlotN20Cluster1}
+       \caption{Cluster 1 Morris Plot, with N =20}
        \label{fig:Cluster1Morris}
 \end{figure}
-
 \begin{figure}[ht]
                 \centering
-        \includegraphics[width=\linewidth]{figures/MorrisPlotN100Cluster2}
-       \caption{Cluster 2 Morris Plot, with N =100}
+        \includegraphics[width=\linewidth]{figures/MorrisPlotN20Cluster2}
+       \caption{Cluster 2 Morris Plot, with N =20}
        \label{fig:Cluster2Morris}
 \end{figure}
-
 %\begin{table}[ht]
@@
 %\end{table}
-
 \begin{table}[ht]
@@
 \end{table}
-
 \begin{table}[]
@@
 \end{table}
-
 %\begin{table}[h]
@@
 %\end{table} 
-
 \begin{table}[h]
@@
 \tiny
-\caption{Average Overall State Sensitivity Coefficents For Single Objective Case}
+\caption{Average Overall State Sensitivity Coefficients For Single Objective Case}
 \label{tab:OSSCsoriginalParams}
